raptr: Representative and Adequate Prioritisation Toolkit in R

### 2016-11-04 code/rmarkdown/supporting-information.Rmd
+ edits to ms + increase blm value

# Appendix S1: Reliable formulation
The reliable formulation explicitly considers the probability that the planning units are inhabited. As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented. This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s). To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.

@@

A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation. Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$. The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$. However, the $d$ variables are in distance units which are meaningless in this context. Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).

@@

Moderately-sized conservation planning problems often include several thousand planning units. It is currently not be feasible to solve this problem when considering all possible failure scenarios. As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level. Cui et al. [-@r16] found that $R=5$ yields similar solutions to $R=J$ when $J >> 5$. However, in most costs the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.

The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{sa}$ (eqn 1c), $R$, and $M$ variables.

@@

The reliable formulation is non-linear. However, the non-linear components can be linearised. First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al. [-@r426]. Second, the expression $P_{fsijr} Y_{jsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al. [-@r16].

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$.',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$.',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$.',
	'$T_f$', 'amount target for feature $f$.',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$.',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0).',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units.'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')

>>>>

## Appendix S1
The reliable formulation explicitly considers the probability that planning units are inhabited by features. As a consequence, it may deliver prioritisations that will sufficiently represent an attribute space even if the features do not inhabit several of the planning units when the prioritisation is implemented. This behaviour is achieved by siting back-up planning units near selected planning units with low occupancy probabilities in the attribute space(s). To ensure that prioritisations are robust against multiple planning units being uninhabited, the problem assigns planning units at multiple backup levels.

@@

A final backup $r$-level when $r=R$ is used to assess the level of representation when the features $f \in F$ do not occupy any selected planning units in a prioritisation. Each demand point $i \in I$ for each $s \in S$ and $f \in F$ is assigned to an "imaginary" planning unit $j=J$ at $r=R$. The distance variables associated with this imaginary planning unit $d_{fsiJ}$ denote the loss of biological value associated with failing to secure a representative sample of feature $f$ in attribute space $s$. However, the $d$ variables are in distance units which effectively meaningless in this context. Thus these variables are calculated using a failure multiplier ($M$) and the maximum distance between the planning units and the demand points for $f \in F$, $s \in S$ (4).

@@

Moderately-sized conservation planning problems often include several thousand planning units. It is currently not be feasible to solve this problem when considering all possible failure scenarios. As a consequence, the $R$ variable can be any $1 \leq R \leq J-1$. For instance, when $R=3$ only 2 backup levels are considered in addition to the final backup level. Cui et al. [-@r16] found that $R=5$ yields similar solutions to $R >> 5$. However, in most cases the decision maker will likely be limited to $R=1$ to obtain prioritisations in a feasible amount of time.

The control variables for the reliable formulation are the $B$ (eqn 1a), $T_{s}$ (eqn 1b), $\tau_{fs}$ (eqn 1c), $R$, and $M$ variables.

@@

The reliable formulation is non-linear. However, the non-linear components can be linearised. First--as discussed in the main text--the expression $X_j X_k$ in (eqn 3a) can be linearised using methods described by Beyer et al. [-@r426]. Second, the expression $P_{fsijr} Y_{fsijr}$ in (eqn 7a) can be linearised using techniques described by Sherali and Alameddine [-@r20] as implemented in Cui et al. [-@r16].

@@

```{r, echo=FALSE, results='asis'}
# make table
term.CHR <- c(
	'$F$', 'set of biodiversity features (indexed by $f$)',
	'$J$', 'set of planning units (indexed by $j$)',
	'$q_{fj}$', 'probability of feature $f$ occupying planning unit $j$',
	'$e_{jk}$', 'shared edge between planning unit $j$ and planning unit $k$. Note that when $j==k$ this used to parametrise exposed edges with no neighbours.',
	'$S$', 'set of attribute spaces (indexed by $S$)',
	'$I_{fsi}$', 'set of demand points (indexed by $i$) for a feature $f$ in attribute space $s$',
	'$\\lambda_{fsi}$', 'set of weights for demand point $i$ for feature $f$ in attribute space $s$',
	'$d_{fsij}$', 'distance between demand point $i$ and planning unit $j$ for feature $f$ in attribute space $s$',
	'$\\delta_{fsi}$', 'the distance between each demand point $i$ and the centroid of the demand points $I$ for feature $f$ in attribute space $s$',
	'$T_f$', 'amount target for feature $f$',
	'$\\tau_{fs}$', 'space-based target for feature $f$ in attribute space $s$',
	'$X_j$', 'binary decision variable controlling if a planning unit is selected for preservation (1) or discarded (0)',
	'$Y_{fsij}$', 'binary decision variable indicating if planning unit $j$ is assigned to demand point $i$ for species $s$ in a attribute space $s$ when determing the amount of the attribute space sampled by the selected planning units'
)
term.DF <- data.frame(
	Symbol=term.CHR[seq(1, length(term.CHR), 2)],
	Description=term.CHR[seq(2, length(term.CHR), 2)]
)
term.DF$Description <- paste0('\\parbox[t][][t]{15cm}{', term.DF$Description, '}')
